Machine Learning for Predictive Auto-Tuning

### 2011-12-12 paper/draft1.tex
Fix formatting issues

%XXX:  Machine Learning for Auto-tuning: A Data-Drive Approach...} ???
%\title{Boosted Regression Trees for Data-Driven Auto-Tuning}
\title{Machine Learning for Predictive Auto-Tuning}
%XXX:  Machine Learning for Auto-tuning: A Data-Drive Approach...} ???

@@

Due to power consumption and heat dissipation concerns, scientific applications
have shifted from computing platforms where performance had been primarily
driven by rises in the clock frequency of a single ‘’heavy-weight’’ processor
(with complex out-of-order control and cache structures) to a platform with
ever increasing numbers of ‘’light-weight’’ cores. Interestingly, this shift is
now not only relevant to computational sciences but to the development of all
computer systems: from ubiquitous consumer-facing devices (e.g. phones) to
high-end computer farms for web-scale applications (e.g. social networks).

@@

% 1.1 Motivation
% Better tools: motivation (compilers/tools/libs can’t keep up, it’s too difficult)

In this rapidly evolving landscape, the construction of \emph{general}
development tools and libraries that \emph{fully utilize} system resources
remains a daunting task \cite{?}. Even within specialized architectures from
the \emph{same} vendor, like NVIDIA’s Graphics Processing Units (GPUs) and the
Compute Unified Device Architecture (CUDA) \cite{CUDA}, many developers default
to massive amounts of manual labor to optimize CUDA code to specific input
domains. In addition, hand-tuning rarely generalizes well to new hardware
generations or different input domains, and it can also be error-prone or far
from optimal. One of the reason is that kernels can produce staggeringly large
optimization spaces \cite{K. Datta et al. Stencil computation optimization and
auto-tuning on state-of-the-art multicore architectures}. The problem is
further compounded by the fact that these spaces can be highly discontinuous
\cite{RyooHwuCGO2008}, difficult to explore and holding quasi-optimal solutions
at the edge of ‘’performance cliffs’’ (in part due to hard resource
restrictions, like the register file size).

@@

One of the strategies for addressing these challenges is to use automatic
methods called ‘’auto-tuning’’. The literature on auto-tuning is extensive (see
surveys in \cite{XXX}. Two different approaches to auto-tuning have emerged:
model-driven optimization and empirical optimization.

>>>>

%XXX:  Machine Learning for Auto-tuning: A Data-Drive Approach...} ???
%\title{Boosted Regression Trees for Data-Driven Auto-Tuning}
\title{Machine Learning for Predictive Auto-Tuning}
% NP: I like this title !
%XXX:  Machine Learning for Auto-tuning: A Data-Drive Approach...} ???

@@

Due to power consumption and heat dissipation concerns, scientific applications
have shifted from computing platforms where performance had been primarily
driven by rises in the clock frequency of a single ''heavy-weight'' processor
(with complex out-of-order control and cache structures) to a platform with
ever increasing numbers of ''light-weight'' cores. Interestingly, this shift is
now not only relevant to computational sciences but to the development of all
computer systems: from ubiquitous consumer-facing devices (e.g. phones) to
high-end computer farms for web-scale applications (e.g. social networks).

@@

% 1.1 Motivation
% Better tools: motivation (compilers/tools/libs can't keep up, it's too difficult)

In this rapidly evolving landscape, the construction of \emph{general}
development tools and libraries that \emph{fully utilize} system resources
remains a daunting task \cite{?}. Even within specialized architectures from
the \emph{same} vendor, like NVIDIA's Graphics Processing Units (GPUs) and the
Compute Unified Device Architecture (CUDA) \cite{CUDA}, many developers default
to massive amounts of manual labor to optimize CUDA code to specific input
domains. In addition, hand-tuning rarely generalizes well to new hardware
generations or different input domains, and it can also be error-prone or far
from optimal. One of the reason is that kernels can produce staggeringly large
optimization spaces \cite{K. Datta et al. Stencil computation optimization and
auto-tuning on state-of-the-art multicore architectures}. The problem is
further compounded by the fact that these spaces can be highly discontinuous
\cite{RyooHwuCGO2008}, difficult to explore and holding quasi-optimal solutions
at the edge of ''performance cliffs'' (in part due to hard resource
restrictions, like the register file size).

@@

One of the strategies for addressing these challenges is to use automatic
methods called ''auto-tuning''. The literature on auto-tuning is extensive (see
surveys in \cite{XXX}. Two different approaches to auto-tuning have emerged:
model-driven optimization and empirical optimization.
